---
title: "R Notebook"
output: html_notebook
---
```{r}
library("Hmisc")
library("tidyr")
library("mice")
library("dplyr")
library("ggplot2")
library("BaylorEdPsych")
library("caret")
library("caretEnsemble")
```



Read in Sets
```{r}
train<-read.csv('/Users/Cam/Documents/Final Project/train.csv', header=TRUE, sep=",", na.strings = c("NA",""))
test<-read.csv('/Users/Cam/Documents/Final Project/test.csv', header=TRUE, sep=",", na.strings = c("NA",""))
str(train)
train_labels<-train[,1:2]
```


```{r}

train$Survived<-NULL
train<-rbind(train,test)
train<-merge(train,train_labels,all=T)
```

Printing the amount of NA in each column
```{r}
na_countTrain<-apply(train, 2, function(x) sum(is.na(x)))
na_countTest<-apply(test, 2, function(x) sum(is.na(x)))
print(na_countTrain)
print(na_countTest)


```

```{r}
keepTrain<-c( "Name", "Sex", "Age", "Pclass", "SibSp", "Parch", "Fare", "Embarked", "Survived")
keepTest<-c("Name", "Sex", "Age", "Pclass", "SibSp", "Parch", "Fare", "Embarked")
train<-train[keepTrain]
test<-test[keepTest]
```

refactoring data
```{r}
train$Pclass<-as.ordered(train$Pclass)
train$Name<-as.character(train$Name)
train$Survived<-as.factor(train$Survived)
```


Analyzing title as that could indicate sicio economic status
```{r}
train2<-separate(train, Name, into=c("Last Name", "Title", "First Name"), sep="[\\.|\\,]", remove=FALSE)
train2$Title<-gsub("\\s+","",train2$Title)


```

```{r}
table(train2$Sex, train2$Title)
```
```{r}
train2$NameLength<-nchar(train2$Name)
```

Get number of people on ticket
```{r}
sameTicket<-as.data.frame(table(train2$Ticket))
colnames(sameTicket)<-c("Ticket","Ticket.Freq")
train2<-merge(train2, sameTicket, by="Ticket",all=TRUE)
```

More preproccessing title
```{r}
train2$Title[train2$Title %in% c("Mlle", "Ms")] <- "Miss"
train2$Title[train2$Title== "Mme"] <- "Mrs"
train2$Title[!(train2$Title %in% c('Master', 'Miss', 'Mr', 'Mrs'))] <- "Rare Title"
```


```{r}
train2$Fsize<-train2$SibSp+train2$Parch+1

data2<-as.data.frame(table(train2$`Last Name`,train2$Fsize))
colnames(data2)<-c("Last Name","Fsize", "freq")
data2<-data2[which(data2$freq!=0),]

```

```{r}
train2$IsAlone<-sapply(train2$Fsize, function(x) ifelse(x==1,1,0))
```

```{r}

train2$Title<-as.factor(train2$Title)
Age<-subset(train2,train2$Age!="NA")



ageImpute <- train(Age~Title+Sex+Parch+SibSp+IsAlone, data=Age, method="gbm", trControl=control, verbose=FALSE)

```

```{r}
ind2<-Age$PassengerId
train2<-train2[order(train2$PassengerId),]
missingAge<-train2[-ind2,]
predictAge<-predict(ageImpute,missingAge)
```

```{r}
missingAge$Age<-predictAge

```

```{r}
train3<-rbind(Age,missingAge)
```


```{r}
train2$Title<-as.factor(train2$Title)
imp2<- aregImpute(~ Age+Parch+SibSp+ Title+Fsize+Ticket.Freq, data = train2, n.impute = 5, nk=0)
imp2$imputed
imp2

```
```{r}
completedata<-as.data.frame(impute.transcan(imp2, imputation=5, data=train2, list.out=TRUE,pr=FALSE, check=FALSE))
train2$Age<-completedata$Age
```

```{r}
train2<-train2[order(train2$PassengerId),]
ind<-test$PassengerId
test2<-train2[ind,]
train2<-train2[-ind,]

```

```{r}

# Example of Bagging algorithms
control <- trainControl(method="repeatedcv", number=10,repeats=3, search = "grid",savePredictions=TRUE, classProbs=TRUE)
seed <- 7
metric <- "Accuracy"
levels(train2$Survived) <- c("diey", "live")
algorithmList <- c('rf', 'glm', 'nnet','knn', 'svmRadial')
models <- caretList(Survived~Fsize+Embarked+Ticket.Freq+Sex+Age+Parch+SibSp+Pclass, data=train2, methodList=algorithmList, metric=metric, trControl=control)

```


```{r}
results <- resamples(models)
summary(results)
dotplot(results)



```


```{r}
# correlation between results
modelCor(results)
stack.glm <- caretStack(models, method="gbm", metric="Accuracy", trControl=control)
```


```{r}
fit.rf <- train(Survived~Fsize+Embarked+Ticket.Freq+Sex+Age+Parch+SibSp+Pclass, data=train2, method="rf", metric=metric, trControl=control)
TrainMale<-subset(train2, train2$Sex=="male")
fit.gbmM <- train(Survived~Fsize+Embarked+Ticket.Freq+Age+Parch+SibSp+Pclass, data=TrainMale, method="gbm", metric=metric, trControl=control, verbose=FALSE)
```


```{r}

fit.gbm <- train(Survived~Fsize+Embarked+Ticket.Freq+Sex+Age+Parch+SibSp+Pclass, data=train2, method="gbm", metric=metric, trControl=control, verbose=FALSE)
```

```{r}
varImp(models)

```

```{r}

#train2$isNA<-as.integer(is.na(train2$Age))
length(train2$isNA)
#train2$Survived<-as.factor(train2$Survived)
train2$Cabin<-gsub('[[:digit:]]+', "",train2$Cabin)
train2$Cabin<-gsub('\\s+.*', "",train2$Cabin)
```




```{r}
train2$Cabin<-as.factor(train2$Cabin)
train2$HasNoCabin<-as.integer(is.na(train2$Cabin))

```




```{r}

#parch with female vs survived
ggplot(subset(train2, train2$Sex=="female" & train2$Age>=18 & train2$Survived!="NA" & train2$Pclass=="3"), aes(x=Parch, fill=Survived))+geom_bar(stat='Count',position = 'dodge')+theme_grey()+geom_label(stat='count', aes(label=..count..))

#SibSp vs female
ggplot(subset(train2, train2$Sex=="female" & train2$Age>=18 & train2$Survived!="NA" & train2$Pclass=="3"), aes(x=SibSp, fill=Survived))+geom_bar(stat='Count',position = 'dodge')+theme_grey()+geom_label(stat='count', aes(label=..count..))



#Pclass 3 female
ggplot(subset(train2, train2$Sex=="female" & train2$Age>=18 & train2$Pclass=="3" & train2$Survived!="NA"), aes(x=Ticket.Freq, fill=Survived))+geom_bar(stat='Count',position = 'dodge')+theme_grey()+geom_label(stat='count', aes(label=..count..))





```


```{r}
test3M<-subset(test3,test3$Sex=="male")
pred3<-predict(fit.gbmM,test3M)
predictionsMale3<-data.frame(test3M$PassengerId,pred3)
predictionsMale3$pred3<-as.numeric(predictionsMale3$pred3)
predictionsMale3$pred3<-sapply(predictionsMale3$pred3,function(x) x=x-1)
```

```{r}
pred<-predict(stack.glm,test2)
predictions<-data.frame(test2$PassengerId,pred)
predictions$pred<-as.numeric(predictions$pred)
predictions$pred<-sapply(predictions$pred,function(x) x=x-1)
colnames(predictions)<-c("PassengerId","Survived")
write.csv(predictions, file = "MyData.csv", row.names = FALSE)
```


```{r}

#Cabin vs survived
ggplot(subset(train2, train2$Survived!="NA"), aes(x=HasNoCabin, fill=Survived))+geom_bar(stat='Count',position = 'dodge')+theme_grey()+geom_label(stat='count', aes(label=..count..))

#Cabin and in P1
ggplot(subset(train2, train2$Pclass=="1"  & train2$Survived!="NA"), aes(x=HasNoCabin, fill=Survived))+geom_bar(stat='Count',position = 'dodge')+theme_grey()+geom_label(stat='count', aes(label=..count..))

ggplot(subset(train2,train2$HasNoCabin=="0" & train2$Sex=="male" & train2$Pclass=="1" & train2$Survived!="NA"), aes(x=Cabin, fill=Survived))+geom_bar(stat='Count',position = 'dodge')+theme_grey()+geom_label(stat='count', aes(label=..count..))
```


```{r}
ggplot(subset(train2,train2$Survived!="NA"  & train2$Sex=="female" & train2$Pclass=="3"), aes(x=Fsize, fill=Survived))+geom_bar(stat='Count',position = 'dodge')+theme_grey()+geom_label(stat='count', aes(label=..count..))

#Ticket.Freq vs survived
ggplot(subset(train2,train2$Survived!="NA" & train2$Sex=="female" & train2$Pclass=="3"), aes(x=Ticket.Freq, fill=Survived))+geom_bar(stat='Count',position = 'dodge')+theme_grey()+geom_label(stat='count', aes(label=..count..))

ggplot(subset(train2,train2$Survived!="NA" & train2$Sex=="female" & train2$Pclass=="3" & train2$Fsize=="1"), aes(x=Ticket.Freq, fill=Survived))+geom_bar(stat='Count',position = 'dodge')+theme_grey()+geom_label(stat='count', aes(label=..count..))




```





```{r}
ggplot(subset(train,train$Sex=="male" & train$Survived!="NA" & train$Pclass=="1" & train$Age>=18), aes(x=Age, fill=Survived)) +
  geom_density(position="identity")


ggplot(subset(train2,train2$Sex=="female" & train2$Survived!="NA" & train2$Pclass=="3" & train2$Age<=18), aes(x=Age, fill=Survived)) +
  geom_density(position="identity")

ggplot(subset(train,train$Sex=="female" & train$Survived!="NA" & train$Pclass=="3"), aes(x=Age, fill=Survived)) +
  geom_histogram(position="identity", bins = 15)

ggplot(subset(train2,train2$Sex=="female" & train2$Survived!="NA" & train2$Pclass=="3"), aes(x=Age, fill=Survived)) +
  geom_histogram(position="identity", bins = 15)

ggplot(subset(train2,train2$Survived!="NA" & train2$Sex=="female" & train2$Pclass=="3"), aes(x=Sex, fill=Survived))+geom_bar(stat='Count',position = 'dodge')+theme_grey()+geom_label(stat='count', aes(label=..count..))

ggplot(subset(train2,train2$Survived!="NA" & train2$Sex=="male"), aes(x=Sex, fill=Survived))+geom_bar(stat='Count',position = 'dodge')+theme_grey()+geom_label(stat='count', aes(label=..count..))
```
```{r}
ggplot(subset(train2,train2$Fsize=="1" & train$Survived!="NA"), aes(x=Age)) +
  geom_density(position="identity")

ggplot(subset(train2,train2$SibSp=="1" & train$Survived!="NA" & train2$Title=="Miss" & train2$Ticket.Freq=="2"), aes(x=Age)) +
  geom_density(position="identity")


```


First I realized that fare is by ticket so one ticket for 2 people is 2x not by person.
```{r}
train2$FarePP<-train2$Fare/train2$Ticket.Freq

```

Imputing missing embarked. Ran t-tests based on my own hypothesis that fare was due to embarked and Pclass. The results back my hypothesis that there is a difference in fare base on embarked and pclass. Now the two missing embarked were from Pclass 1.Also, there are only two entries for Q and Pclass equals 1 so it cannot be from class Q. Now given the results I can say that it is most likely in class S
```{r}

t.test(subset(train2$FarePP,train2$Embarked=="S" & train2$Pclass=="1" ),subset(train2$FarePP,train2$Embarked=="C" & train2$Pclass=="1"))

t.test(subset(train2$FarePP,train2$Embarked=="C" & train2$Pclass=="1"))

t.test(subset(train2$FarePP,train2$Embarked=="S" & train2$Pclass=="1"))








```
```{r}
train2$Embarked[train2$Ticket=='113572']<-"S"
```


```{r}
ggplot(train2, aes(x = Embarked, y = Fare, fill=Pclass )) +
    geom_boxplot() + scale_y_continuous(breaks = scales::pretty_breaks(n = 10)) + theme_grey()


ggplot(train2, aes(x = Embarked, y = Fare)) +
    geom_boxplot() + scale_y_continuous(breaks = scales::pretty_breaks(n = 10)) + theme_grey()
```


```{r}

```


```{r}
analyze<-function(x){
  chisq.test(table(train2$isNA,x))}

```
Assuming not Missing Not At Random this test is to see whether its Missing Completely at Random or Missing At random as a result we can tell that in every case for p-value=.05 we reject the null hypothesis that they are missing competely at random. 

```{r}

TestForMissingAtRandom<-lapply(train2, function(x) if(is.integer(x)){analyze(x)})
print(TestForMissingAtRandom)
```
```{r}
imp<-mice(train2[c(-1,-2,-4,-12,-13)],m=5, method='pmm', seed=5565)
data2<-complete(imp,action='all')
fit <- with(data = imp, lm(Age~Title+Sex+Pclass+Embarked+SibSp+Parch+Fare))
summary(pool(fit))

```

